\cleardoublepage 

# Comparison--barchart and boxplots{#Comparison}




## Graph considerations for communication: aggregation, abstraction, complexity 
### Simple bar chart
```{r message=FALSE}
library(tidyr)
library(ggforce)
library(ggthemes)
library(ggstance)

mtcars.df = mtcars
mtcars.df = mtcars.df %>% mutate(cyl = as.factor(cyl))

s.mtcars.df = mtcars.df %>% group_by(cyl) %>% summarise(m.hp = mean(hp), se.hp= sd(hp)/n()^.5)


ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_bar(stat="identity")

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_col()

## Change order of bars
cyl.order <- c("8", "6", "4")
ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_col() +
  scale_x_discrete(limits = cyl.order)
```


### Bar chart with error bars
```{r}

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_bar(stat="identity")+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_bar(stat="identity")+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))+
  geom_point(data = mtcars.df, aes(cyl, hp), position = position_jitter(width = .2, height = 0))
```



### dotplot and offset range plot
```{r, cache=TRUE}
## Set seed and create data
set.seed(999)
df = data_frame(A = runif(12,1,17), B = runif(12, 2, 8))


l.df = gather(df, condition, value)
l.df$condition = as.factor(l.df$condition)
m.l.df = l.df %>% group_by(condition) %>% summarise(m.value = mean(value, na.rm=TRUE), 
      n= sum(!is.na(value)), sd=sd(value, na.rm=TRUE), sde=sd(value, na.rm=TRUE)/n^.5,
       ci= 2*sde)
m.l.df$n.condition = as.numeric(m.l.df$condition)-.05 


## Plot with offset for mean and error bar
ggplot()+
  geom_dotplot(data = filter(l.df, condition=="A"|condition=="B"),
               aes(condition, value), binaxis = "y", stackdir = "up")+
  geom_linerange(data = filter(m.l.df, condition=="A"|condition=="B"), 
                 aes(n.condition, ymin=m.value-ci, ymax=m.value+ci), color="grey50") +
  geom_point(data = filter(m.l.df, condition=="A"|condition=="B"), 
             aes(n.condition, y= m.value), shape = 21, size = 4, fill="grey", alpha=.7) +
  labs(x="", y="") +
  ylim(2, 15)
```      


### Statistical significance in context
```{r}
ggplot(data = mtcars.df, aes(x = as.factor(cyl), y = hp)) + 
  geom_boxplot(colour = "darkgrey") + 
  geom_point(stat="summary", fun.y = "mean", size = 6, shape = 1) +
  geom_pointrange(stat="summary", fun.data = "mean_cl_boot") +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 1, 
               dotsize = 6,alpha = .3, color = "black", fill = "red") +
  geom_hline(aes(yintercept = mean(hp)), size = 1.2) 
```
  


## Comparing distributions, box, violyn, and sina plot
```{r message=FALSE}
library(tidyr)

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_col()+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_col()+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))+
  geom_point(data = mtcars.df, aes(cyl, hp), position = position_jitter(width = .2, height = 0))

ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_point(stat="identity", size = 3)+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))+
  geom_point(data = mtcars.df, aes(cyl, hp), position = position_jitter(width = .2, height = 0))


## Sina plot
ggplot(mpg, aes(as.factor(cyl), hwy))+
 geom_sina(aes(color = as.factor(cyl)),size = 1, alpha =.5) +
 geom_tufteboxplot()+
  labs(title = "ggforce: sina plot with Tufte boxplot")


# ggplot(s.mtcars.df, aes(m.hp, m.mpg, colour =as.factor(cyl)))+
# geom_pointrangeh(aes(xmin= m.hp-se.hp, xmax = m.hp+se.hp))+
#   geom_pointrange(aes(ymin= m.mpg-se.mpg, ymax = m.mpg+se.mpg))+
#   labs(title = "ggstance: horizontal point range")

ggplot(mtcars.df, aes(as.factor(cyl), hp))+
geom_boxplot()


ggplot(mtcars.df, aes(as.factor(cyl), hp))+
geom_violin(draw_quantiles = c(0.25, 0.5, 0.75))


ggplot(data = s.mtcars.df, aes(x = cyl, y = m.hp)) +
  geom_violin(data= mtcars.df, aes(cyl, hp))+
  geom_point(stat="identity", size = 3)+
  geom_linerange(aes(ymin=m.hp-2*se.hp, ymax=m.hp+2*se.hp))+
  geom_point(data = mtcars.df, aes(cyl, hp), 
             position = position_jitter(width = .2, height = 0), alpha =.6)
```


  

### Compare  empirical and theoretical distribution

```{r}
sum.mtcars.df = mtcars.df%>% group_by(cyl) %>% 
	summarise(m.hp = mean(hp), sd.hp = sd(hp))

ggplot(mtcars.df) +
  geom_boxplot(aes(as.factor(cyl), hp)) +
  geom_linerange(data = sum.mtcars.df,
        aes(x = as.factor(cyl), 
		ymin = m.hp + qnorm(.25)*sd.hp, ymax = m.hp + qnorm(.75)*sd.hp),
                 size = 5, alpha = .25) +
  geom_point(data = sum.mtcars.df,
         aes(as.factor(cyl), y= m.hp),size = 6, alpha = .33)

```
  
  
### Tufte-inspired minimal bar chart
http://motioninsocial.com/tufte/

```{r}
#TODO replace with better dataset with for more columns
library(ggthemes)
ggplot(mtcars.df, aes(x=as.factor(cyl))) + 
  geom_bar(width=0.25, fill="gray") +  
  scale_y_continuous(breaks=seq(2, 12, 2)) + 
  geom_hline(yintercept=seq(2, 12, 2), colour="white", lwd=1) +
  theme_tufte(base_size=12, ticks=F, base_family = "Arial") 
```
   

## Comparing across many variables
### Dot plots and reordering
Comparing many mean values
```{r}
#TODO Change to mean sd catepilar plot
mtcars.df = mtcars
mtcars.df$name = rownames(mtcars.df)
ggplot(data = mtcars.df, aes(reorder(name, hp), y = hp, colour = as.factor(cyl))) +
  geom_point(size = 2) +
  geom_hline(aes(yintercept = mean(hp)), colour = "darkgrey") +
  geom_linerange(aes(ymin= -Inf, ymax= hp), alpha =.5) +
  coord_flip() +
  labs(x = "Cylinders", y = "Power (hp)") +
  scale_colour_brewer(name = "Number of \nCylinders", palette="Dark2") + # http://colorbrewer2.org
  theme(legend.position = c(.75, .25)) +
  theme_minimal()
```



### Point range on x and y
```{r, xycomparison}
## Point range on x and y
library(ggstance)
s.mtcars.df = mtcars %>% group_by(cyl) %>% 
  summarise(m.hp = mean(hp), se.hp = sd(hp)/n()^.5,
            m.mpg = mean(mpg), se.mpg = sd(mpg)/n()^.5)
```

    
### Tufte boxplot for many variables
```{r, cache=TRUE}
library(ggthemes)

ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_tufteboxplot(median.type = "line", whisker.type = 'line', hoffset = 0, width = 4) +
  geom_rangeframe() 

## Tufte boxplot
ggplot(mtcars, aes(as.factor(cyl), mpg))+
  geom_tufteboxplot()+
  labs(title = "ggthemes: Tufte boxplot")

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl)))+
geom_point()+
  geom_rangeframe(size = 2, colour = "grey35")+
  labs(title = "ggthemes: Tufte range frame")

```

### Tufte-inspired slope graphs {#tufte-slopegraph}
```{r, fig.asp = 1.3, fig.height = 5}
library(tidyverse)
library(ggrepel)

# https://github.com/leeper/slopegraph
cancer.df = read_csv("data/tufte-cancer-survival-data.csv")

l.cancer.df = cancer.df %>% gather(key = year, value = rate, 2:5)

l.cancer.df$year = factor(l.cancer.df$year, 
                          levels = c("Year 5", "Year 10", "Year 15", "Year 20"))

ggplot(l.cancer.df, aes(year, rate, group = Type))+
  geom_line(colour = "grey70") +
  geom_text_repel(data = l.cancer.df %>% filter(year == "Year 5"),
                  aes(label = Type), nudge_x = -.35, direction = "y",
                  point.padding = .02)+
  geom_text_repel(data = l.cancer.df %>% filter(year=="Year 20"),
                  aes(label = Type), nudge_x = .35,  direction = "y",
                  point.padding = .02)+
  geom_label(aes(label = rate), colour = "grey55", label.size = .02)+
  theme_void()+
  theme(axis.text = element_text(size = rel(.85)),
        axis.text.y=element_blank())
  
```

### Parallel coordinate plot with similar items highlighted
Link to network for showing links in multivariate data

```{r, cache=TRUE, echo=FALSE, warning=FALSE}
## Highlight closest pair of items multidimensional space ##
library(tidyr)

mtcars.df = mtcars
mtcars.df$name = row.names(mtcars.df)


## Scale variables and calculate distances between cases
mtcars.df = mtcars.df %>% mutate(gpm = 1/mpg) %>% mutate(speed = 1/qsec) %>% 
  dplyr::select(-mpg, -qsec) %>% 
  dplyr::select(cyl:carb, gpm, speed, name)

scaled_mtcars.df = mtcars.df %>% mutate_at(vars(cyl:speed), scale)

dist.df =dist(scaled_mtcars.df, upper = TRUE, diag = FALSE)
dist.df = as.data.frame(as.matrix(dist.df))


## Find the closest vehicle to target vehicle
target = 18
closest = which(min(dist.df[dist.df[, target]>0, target])==dist.df[, target])


## Convert to long format for parallel coordinate plot
l.scaled_mtcars.df = scaled_mtcars.df %>% gather(var, value, -name)

mtcars.df$name[target]
mtcars.df$name[closest]

pair = l.scaled_mtcars.df %>% filter(name==mtcars.df$name[target]|name==mtcars.df$name[closest])

ggplot(l.scaled_mtcars.df, aes(reorder(var, -value, sd), value, group = name))+
  geom_line(alpha = .3, size = .2)+
  geom_point(data = pair, colour = "red", alpha = .5) +
  geom_line(data = pair, colour = "red", alpha = .5) +
  theme_minimal()
```


## Gliphs: Chernof face and radar plots 
Show patterns and outliers not precise comparisons


