\cleardoublepage 

# Highlighting, annotating, polishing, and automating graphs{#Polishing}


```{r, message=FALSE}
library(gapminder)
library(tidyverse)
library(ggrepel)
library(hrbrthemes)
library(ggpubr)
rm(list=ls())

mtcars.df = mtcars

```

All graphs should have axis labels with units
Legends should have titles
Titles are often included in the figure caption and not the graph
Annotations can label points and tell a story
Annotations can label lines better than legends

Labels and annotations added as another layer

Create a plot
Add axis labels, with units
Annotate to tell a story
Select and appropriate color palette
Place the legend inside the graph
Change the theme to increase the data to ink ratio
Check that changing the theme did not undo your legend placement
Save as a 5X5 inch image in a format to minimize blur


## Highlighting datapoints to make a point

```{r}
h.mtcars.df = mtcars.df %>% mutate(highlight = if_else(cyl==8, "yes", "no"))
ggplot(h.mtcars.df, aes(x = as.factor(cyl), fill = highlight)) +
    geom_bar(alpha = .6) +
    scale_fill_manual(values = c("yes"="darkred", "no"="gray15"), guide = FALSE)


h.mtcars.df = mtcars.df %>% mutate(highlight = if_else(mpg>24 &wt>3, "yes", "no"))
ggplot(h.mtcars.df, aes(x = wt, y = mpg, colour = highlight)) +
    geom_point() +
    scale_color_manual(values = c("yes"="darkred", "no"="gray60"), guide = FALSE)
    
    
```


    

## Labeling and annotation
### Annotating data
```{r, annotation}
## Text and rectangle annotation 
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) + 
  annotate(geom = "rect", ymin = 20, ymax = 35, xmin = 65, xmax = 155, fill = "grey65", alpha = .5) +
  annotate(geom = "text", x = 115, y = 36, label = "High-power cars" ) +
  geom_point()+
  labs(x = "Displacement (cubic inches)", y = "Efficiency (mile per gallon)", 
       title = "Efficiency and engine size", colour = "Number of cylinders")


## Encircle pooints
library(ggalt)

 ggplot(mtcars, aes(disp, mpg))+
    geom_point()+
    geom_encircle(data=filter(mtcars, mpg>24),
                   s_shape=0.75, expand=0.05) +
   labs(title = "ggalt: encircle annotation")


## Regresion equation
library(ggpmisc) # For stat_poly_eq for equation annotation  
# https://cran.r-project.org/web/packages/ggpmisc/vignettes/user-guide-1.html

formula <- y ~ poly(x, degree = 1, raw = TRUE)

ggplot(mtcars, aes(x=disp, y=mpg)) + 
    geom_point() +
  geom_smooth(method = "lm", formula = formula) +
  stat_poly_eq(aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
                   formula = formula, parse = TRUE, coef.digits = 2, 
                   label.x = 200, label.y = 30)


```


### Time series with lines labeled
Based on example from: https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html

```{r, line-annotation}
library(ggrepel)

```


### Labelling lines, bars, points, and select points

```{r, point-line-label}
library(ggrepel)

mtcars.df = mtcars
mtcars.df$name = row.names(mtcars.df)


## Lines with labels rather than legend
ggplot(Orange, aes(age, circumference, color = Tree)) +
  geom_line() +
  geom_text_repel(data = filter(Orange, age == max(age)),
    aes(label = paste("Tree", Tree)),
    size = 4, nudge_x = 45, segment.color = NA
  ) +
  coord_cartesian(xlim = c(min(Orange$age), max(Orange$age) + 90)) +
  theme(legend.position = "none") +
  labs(x = "Age (days)", y = "Circumference (mm)")

count.mtcars.df = mtcars.df %>% 
  mutate(cyl = as.factor(cyl), am = as.factor(am)) %>% 
  group_by(am, cyl) %>% 
  summarise(count = n()) %>% 
  mutate(label.pos = cumsum(count) - (0.1 * count))

ggplot(count.mtcars.df, aes(x = cyl, y = count, fill = am, label = count)) +
  geom_bar(stat = "identity") +
  geom_text(size = 3, position = position_stack(vjust = 0.5))


## Annotated point
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) +   
  geom_point()+  
  geom_text_repel(aes(label = name), size = 2, show.legend = FALSE)


## Subset annotated
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) +   
  geom_point()+  
  geom_text_repel(data = mtcars.df %>% filter(mpg>30), aes(label = name), size = 2, show.legend = FALSE)


```

### Meaningful symbols
```{r}
ggplot(mtcars, 
		aes(disp,  mpg, colour = as.factor(am), shape = as.factor(am))) +
  geom_point(size = 8) +   
	scale_color_manual(values = c("0" = "blue", "1" = "red")) +  
  scale_shape_manual(values = c("0" = "\u2642", "1" = "\u2640")) +  
  labs(title = "custom shape: Male/Female")

```


### Annotating with an inset plot
```{r, inset-plot, cache=TRUE}
library(tidyverse)
library(ggforce)
diamonds.df = diamonds

## Specify area of the plot to highlight
xmin = .4; xmax = .75
ymin = 800; ymax = 4000

higlight.df = diamonds.df %>% 
  filter(carat > xmin & carat < xmax, 
         price > ymin & price < ymax)


## Create overall plot
overall.plot = ggplot(diamonds.df, aes(carat, price)) +
  geom_point(alpha = .3, size = .3) +
  annotate(geom = "rect", xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
            alpha = .3, fill = "tomato", color = "tomato") 
  # Use annotate rather than geom_rect otherwise rectangle is plotted for every datapoint

## Create inset plot
inset.plot = ggplot(higlight.df, aes(carat, price)) +
  geom_point(alpha = .3, size = .4, shape = 21, stroke = .2) + 
  labs(x = "", y = "") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "#f6d2cc", color = "tomato"))

## Create custom annotation and add to annotate plot
inset.grob <- ggplotGrob(inset.plot )

combined.plot = overall.plot + 
    annotation_custom(grob = inset.grob, 
                       xmin = 3.5, xmax = 5.2, ymin = 250, ymax = 13000) +
  geom_link0(x = xmax, y = ymax, xend = 3.5, yend = 13000, colour = "tomato", alpha = .3, size =.3) +
  geom_link0(x = xmax, y = ymin, xend = 3.5, yend = 250, colour = "tomato", alpha = .3, size = .3)
  

combined.plot +
  labs(title = "Bigger diamonds cost more", 
       subtitle = "Odd gap for prices around $1500",
       caption = "data from tidyverse \"diamonds\" ",
       x = "Diamond size (carats)", y = "Price ($US)") +
  theme_minimal()
  
```


### Labels on lines
```{r}

library(ggpmisc) # For stat_poly_eq for equation annotation  
formula <- y ~ poly(x, 1, raw = TRUE)
ggplot(mtcars, aes(x=disp, y=mpg)) + 
	geom_point() +  
	geom_smooth(method = "lm", formula = formula) +  	stat_poly_eq(aes(label =  paste(..eq.label.., ..adj.rr.label.., 
		sep = "~~~~")), formula = formula, parse = TRUE, coef.digits = 2, label.x = 200, label.y = 30)

```


### Adding and removing labels: title, axis labels, and facet labels {#AddingRemovingLabels}
```{r}
library(tidyverse)
library(ggalt)
library(ggrepel)

# Add names of cars to dataframe
mtcars.df = mtcars
mtcars.df$name = row.names(mtcars.df)

## Axis labels, title, and legend
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) + 
  geom_point()+
  labs(x = "Displacement (cubic inches)", y = "Efficiency (mile per gallon)", 
       title = "Efficiency and engine size", colour = "Number of cylinders")


## Subtitle and caption
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) + 
  geom_point()+
  labs(x = "Displacement (cubic inches)", y = quote(alpha + beta + frac(delta, theta)), 
       title = "Efficiency and engine size", subtitle = "mtcars data",
       caption = "Caption if needed",
       colour = "Number of cylinders")


## Remove labels 
ggplot(mtcars.df, aes(disp,  mpg, color = as.factor(cyl))) + 
  geom_point()+
  labs(x = "", y = "")


## Facet labels 
cyl_names <- as_labeller(c("4" = "Four", "6" = "Six", "8" = "Eight"))

 ggplot(mtcars.df, aes(wt, mpg))+ 
   geom_point() +
   facet_grid(cyl ~ ., labeller = as_labeller(cyl_names))
 
```


## Position legend
```{r, legend}

# Place legend at bottom
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point() +
  theme(legend.position = "bottom")

# Place legend in graph
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point() +
  theme(legend.position = c(0.9, 0.8))

```
 
### Remove one or more legends
```{r, remove-legend}

ggplot(mtcars.df, aes(wt, mpg, colour = cyl, size = hp))+
  geom_point()

## Remove legend for single mapping
ggplot(mtcars.df, aes(wt, mpg, colour = cyl, size = hp))+
  geom_point()+
  guides(colour=FALSE)

ggplot(mtcars.df, aes(wt, mpg, colour = cyl, size = hp))+
  geom_point()+
  scale_colour_continuous(guide=FALSE) 

## Remove legend for a single layer
ggplot(mtcars.df, aes(wt, mpg, colour = cyl, size = hp))+
  geom_point(show.legend=FALSE)

# Remove all legends
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point() +
  theme(legend.position = "none")

```



### Setting limits on axis and position of graph

 ```{r, plot-margin}
 
 ## Change axis limits to focus on part of data
ggplot(mtcars.df, aes(wt, mpg))+
   geom_point()
 
ggplot(mtcars.df, aes(wt, mpg))+
 geom_point()+
 lims(x = c(3, 4))
 
ggplot(mtcars.df, aes(wt, mpg))+
 geom_point()+
 scale_x_continuous(limits = c(3, 4))
 
 ```

### Adjust margin between points and edge of plot area
```{r}

ggplot(mtcars.df, aes(wt, mpg))+
 geom_point() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) 

```







## Color
The palette of colors that is scaled to the data is set separately for "fill" and for "colour"

Color palettes that work for discrete variables (e.g., factors) may not work for continuous, numeric variables

Color palette design is complex and choice depends on

Number of categories in the data
Ability to support black and white printing

Ability to support color-deficient vision

Size of the space being colored

Small areas, such as points, benefit from saturated color, and large areas, such a bars are better with less intense color.

Great resources to explain the basis of the Brewer, Viridis, and ptol palettes: http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3

https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html

https://personal.sron.nl/~pault/

 

### Resources to explain the basis form color choice
Brewer, Viridis, and ptol palettes
http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3
https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
https://personal.sron.nl/~pault/


### Large and small area colors
```{r, cache=TRUE, echo=FALSE}
library(hrbrthemes) # Themes tuned for font 
library(ggthemes) # Many themes including tufte an ptol colors
library(viridis)
library(RColorBrewer)

lc1.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
  scale_fill_hue()+
  labs(title = "ggplot: Default")

lc2.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
  scale_fill_brewer(palette = "Accent") +
  labs(title = "Brewer: Accent")

lc3.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
 scale_fill_brewer(palette = "Dark2") +
 labs(title = "Brewer: Dark2")

lc4.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Brewer: Set2")

## hrbrthemes
lc5.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
  scale_fill_ipsum() +
  labs(title = "hrbrthemes: ipsum")
  
lc6.plot = ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
  scale_fill_ptol() +
  labs(title = "ggthemes: ptol")
  
## viridis
# ggplot(diamonds, aes(x = color, fill = cut)) + geom_bar() +
#   scale_fill_viridis(discrete = TRUE) +
#   labs(title = "Viridis: viridis")

large.color.plot = ggarrange(lc1.plot, lc2.plot, lc3.plot, 
                             lc4.plot, lc5.plot, lc6.plot,
                       nrow=3, ncol = 2, align = "hv")
large.color.plot
```


### Color for large and small areas
What works for large areas of color might not work for small areas
```{r, , cache=TRUE, echo = FALSE}
c1.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+
  scale_colour_viridis(discrete = TRUE)+
  labs(title = "Viridis: viridis")

c2.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+
  scale_colour_brewer(palette = "Dark2")+
  labs(title = "Brewer: Dark2")

c3.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+
  scale_colour_brewer(palette = "Set1")+
  labs(title = "Brewer: Set1")

c4.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+
  scale_colour_brewer(palette = "Set2")+
  labs(title = "Brewer: Set2")

## hrbrthemes
c5.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+ 
  scale_colour_ipsum()+
  labs(title = "hbrbthemes: ipsum")

c6.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl)))+geom_point()+ 
  scale_color_ptol()+
  labs(title = "hbrbthemes: ptol")

small.color.plot = ggarrange(c1.plot, c2.plot, c3.plot, c4.plot, c5.plot, c6.plot,
                       nrow=3, ncol = 2, align = "hv")
small.color.plot
```


### Continuous color scales
```{r}
# Scales made for factors do not work with continuous variables
library(viridis)
ggplot(mtcars, aes(disp,  mpg, color = mpg))+geom_point()+
  labs(title = "gplot: Default")
  
ggplot(mtcars, aes(disp,  mpg, color = mpg))+geom_point()+ 
  scale_color_viridis(discrete = FALSE) +
  labs(title = "viridis: viridis")

##TODO Add parula

```

### Discrete color mappings
```{r}
ggplot(diamonds, aes(x = color, fill = cut)) + 
	geom_bar() +
  scale_fill_hue()+ 
  labs(title = "ggplot: Default")

ggplot(diamonds, aes(x = color, fill = cut)) + 	geom_bar() + 
  scale_fill_brewer(palette = "Accent") + 
  labs(title = "Brewer: Accent")

```








## Themes and theme options

Turn off many theme elements
```{r}
ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
geom_point() +
	theme(axis.line=element_blank(), axis.text.x=element_blank(), 	axis.text.y=element_blank(), axis.ticks=element_blank(), 	axis.title.x=element_blank(), axis.title.y=element_blank(), 	legend.position="none", panel.background=element_blank(), 	panel.border=element_blank(), panel.grid.major=element_blank(), 	panel.grid.minor=element_blank(), plot.background=element_blank())

```


### Remove chart details 
Useful for plotting graphs and networks and maps
```{r}
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point() +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
  
```


### Predefined themes adjust many elements

```{r}
library(hrbrthemes) # Precise font and minimal grid
library(ggthemes) # Huge variety of themes including Tufte and Few
library(xkcd) # Plots in the xkcd comic style
```


An engaging and fun theme
ftp://200.236.31.7/CRAN/web/packages/xkcd/vignettes/xkcd-intro.pdf


```{r, xkcd-plot, cache=TRUE, warning=FALSE, echo=FALSE}
library(extrafont)

xrange <- range(mtcars$wt)
yrange <- range(mtcars$mpg)

p =  ggplot(mtcars.df) + geom_point(aes(wt, mpg)) +
     theme_xkcd()+
     xkcdaxis(xrange, yrange)
 

## Add man
 ratioxy <- diff(xrange) / diff(yrange)

mapping <- aes(x=x,
               y=y,
               scale=scale,
               ratioxy=ratioxy,
               angleofspine = angleofspine,
               anglerighthumerus = anglerighthumerus,
               anglelefthumerus = anglelefthumerus,
               anglerightradius = anglerightradius,
               angleleftradius = angleleftradius,
               anglerightleg =  anglerightleg,
               angleleftleg = angleleftleg,
               angleofneck = angleofneck,
               color = color )

dataman <- data.frame( x= c(2), y=c(23),
                  scale = c(5),
                  ratioxy = ratioxy,
                  angleofspine =  - pi / 2 ,
                  anglerighthumerus = -pi/6,
                  anglelefthumerus = pi + pi/6,
                  anglerightradius = 0,
                  angleleftradius = - pi/4,
                  angleleftleg = 3*pi/2  + pi / 12 ,
                  anglerightleg = 3*pi/2  - pi / 12,
                  angleofneck = -pi / 2 - pi/10,
                  color = "red")

p + xkcdman(mapping, dataman)

```




```{r}
ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
	geom_point() +
	theme_void()
```


```{r}
ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
	geom_point() +
  theme_minimal(base_size = 18)
#, base_family = "Avenir"
```


```{r}
#ipsum, latin for neat

library(hrbrthemes)

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
  geom_point() + 	
  scale_colour_ipsum() + 
	theme_ipsum()

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
  geom_point() + 
  scale_color_fivethirtyeight() +
	theme_fivethirtyeight()

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 
  geom_point()+ 	
  scale_color_few() +
	theme_few()

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + 	
  geom_point()+ 
	geom_rangeframe()+ 
	theme_tufte()

```

order of application matters
```{r}
ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + geom_point()+ 	scale_colour_ipsum() + 
theme_ipsum() +
theme(legend.position = c(.85, .8))

ggplot(mtcars, aes(disp, mpg, color = as.factor(cyl))) + geom_point()+ 	scale_colour_ipsum() + 
theme_ipsum() +
theme(legend.position = c(.85, .8))+  
theme_ipsum()

```








```{r, adjust-text-postion}
ggplot(mtcars.df, aes(wt, mpg)) + 
  geom_point()+
  theme(axis.title.y = element_text(margin = margin(t = 10, r = 10, b = 80, l = 20)))
```






### Ordering theme layers
```{r}
# Sets theme then adjusts legend
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+ 
  theme_ipsum()+
  theme(legend.position = c(.85, .8))

# Adjusts legend, but the overrides with setting theme 
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+ 
  theme(legend.position = c(.85, .8))+
  theme_ipsum()


## To set for all plots
theme_set(theme_few())
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()

theme_set(theme_gray()) # Returns to default theme

```


### Pre-set theme options
```{r, echo=FALSE}
# library(ggpubr)
# 
# gt1.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + 
#   geom_point() +
#   theme(legend.position = "none") +
#   labs(title = "Default: theme_gray()")
# 
# gt2.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + 
#   geom_point() +
#   theme_bw()+
#   theme(legend.position = "none") +
#   labs(title = "theme_bw()")
# 
# # gt3.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + 
# #   geom_point() +
# #   theme_base() +
# #   theme(legend.position = "none") +
# #   labs(title = "theme_base()")
# 
# gt4.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + 
#   geom_point() +
#   theme_minimal() +
#   theme(legend.position = "none") +
#   labs(title = "theme_minimal()")
# 
# # gt5.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point() +
# #   theme_classic2() +
# #   theme(legend.position = "none") +
# #   labs(title = "theme_classic2")
# 
# gt6.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + 
#   geom_point() +
#   theme_void() +
#   theme(legend.position = "none") +
#   labs(title = "theme_void()")
# 
# gtheme.plot = ggarrange(gt1.plot, gt2.plot, 
#                         #gt3.plot, 
#                         gt4.plot, 
#                         #gt5.plot,
#                         gt6.plot,
#                         nrow=3, ncol = 2, align = "hv")
# gtheme.plot
```


### Themes form other packages
```{r, echo=FALSE}
# library(ggpubr)
# ## hbrtthemes
# t1.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+ 
#   scale_colour_ipsum() +
#   theme_ipsum() +
#   labs(title = "hbrthemes: theme_ipsum()")
# 
# ## ggthemes
# t2.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+
#   scale_color_fivethirtyeight() +
#   theme_fivethirtyeight() +
#   labs(title = "ggthemes: theme_fivethirtyeight()")
# 
# t3.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+
#   #scale_color_few() +
#   theme_few() +
#   labs(title = "ggthemes: theme_few()")
# 
# t4.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+
#   #geom_rangeframe() +
#   theme_tufte() +
#   labs(title = "ggthemes: theme_tufte()")
# 
# t5.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+
#   theme_economist_white() +
#   labs(title = "ggthemes: theme_economist_white()")
# 
# t6.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()+
#   theme_excel() +
#   labs(title = "ggthemes: theme_excel()")
# 
# theme.plot = ggarrange(t1.plot, t2.plot, t3.plot, 
#                        t4.plot, t5.plot, t6.plot,
#                        nrow=3, ncol = 2, align = "hv")
# theme.plot

```




## Saving and printing plots to include in documents

For formal documents save graphs and import. Do not cut and paste from RStudio 
Saving and importing provides consistent physical size, resolution, and aspect ratio: DO NOT re-scale in the document
Vector formats (PDF, SVG) 

Provide crisp images even when zoomed in and raster
File size scales with number of data points
Raster formats (PNG, TIFF)

The dpi (dots per inch) defines the resolution of the image	File size scales with dimensions of graph and dpi

### PNG, JPEG, PDF, and SVG 
```{r}
mpg.plot = ggplot(mtcars, 
	aes(disp,  mpg, color = as.factor(cyl))) + 
	geom_point()

ggsave(filename = "mpg.png", 
	device = "png",
	plot = mpg.plot, height = 4, width = 5, units = "in", 
	dpi = 300)

ggsave(filename = "mpg.pdf", device = "pdf", 
      plot = mpg.plot, height = 4, width = 5, units = "in")
```

```{r, cache = TRUE}
library(svglite)
ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()

mpg.plot = ggplot(mtcars, aes(disp,  mpg, color = as.factor(cyl))) + geom_point()

## Possible formats:
#  "eps", "ps", "tex" (pictex), "pdf", "jpeg", "tiff", "png", "bmp", "svg" or "wmf" (windows only)

## PNG--raster format that looks blurry at low resolution  
ggsave(filename = "mpg.png", device = "png",
       plot = mpg.plot, height = 4, width = 5, units = "in", dpi = 300)

## PDF--vector format that remains sharp even when zoomed in
ggsave(filename = "mpg.pdf", device = "pdf",
       plot = mpg.plot, height = 4, width = 5, units = "in")

## SVG--vector format that remains sharp even when zoomed in
ggsave(filename = "mpg.svg", device = "svg",
       plot = mpg.plot, height = 4, width = 5, units = "in")


## Saving many plots into a single file
# Calculate the number of pages with 9 panels per page
# n_pages <- ceiling(
#   length(levels(diamonds$color)) * length(levels(diamonds$cut:diamonds$clarity)) / 9
# )
# 
# pdf("multipage.pdf")
#   for (i in seq_len(n_pages)) {
#     p=  ggplot(diamonds, aes(carat, price)) +
#           geom_point(alpha = 0.1) +
#           facet_grid_paginate(color~cut:clarity, ncol = 3, nrow = 3, page = i)+
#         labs(title = "ggforce: facet pagination")
#     print(p)
# }
# dev.off()

```


### Combining multiple graphs for publication
```{r}
library(ggpubr)

displ.plot = ggplot(mtcars, aes(disp, mpg)) + geom_point()
hp.plot = ggplot(mtcars, aes(hp, mpg)) + geom_point()

combined.plot = ggarrange(displ.plot, hp.plot, nrow=1, ncol = 2, align = "hv")
combined.plot

```


### Faceted pagination
```{r, cache=TRUE}
library(ggforce)
## Examples from: https://cran.r-project.org/web/packages/ggforce/vignettes/Visual_Guide.html


ggplot(diamonds, aes(carat, price)) +
  geom_point(alpha = 0.1) +
  facet_grid_paginate(color~cut:clarity, ncol = 3, nrow = 3, page = 1)+
  labs(title = "ggforce: facet pagination")

```

## Functions with dplyr 
```{r}

gapminder.df = gapminder

mean_grouped<- function(df, group_var, summary_var) {
  group_var = enquo(group_var)
  summary_var = enquo(summary_var)
  df %>% group_by(UQ(group_var)) %>% 
    summarise(m.pop = mean(UQ(summary_var)))
}

mean_grouped(gapminder, continent, pop)
```


## Create functions for custom plot
```{r}
library(gapminder)
library(ggpubr)


timeline_highlight <- function(df, x_var, y_var, size_var, highlight_country) {
  
  ggplot(df, aes_string(x_var, y_var, size = size_var)) +
    geom_point(alpha = .1) +
    geom_point(data = df %>% filter(highlight_country==country), 
               colour = "red") +
    labs(title = highlight_country)
}


US.plot = timeline_highlight(gapminder.df, "year", "lifeExp", "gdpPercap", "United States")
Rwanda.plot = timeline_highlight(gapminder.df, "year", "lifeExp", "gdpPercap", "Rwanda")

combined.plot = ggarrange(US.plot, Rwanda.plot, 
                          nrow=1, ncol = 2, align = "hv", 
                          common.legend = TRUE)
combined.plot
```
