\cleardoublepage 

# Interaction--annotate, highlight, link, and animate{#Interaction}

Animations tend to be less effective than well-crafted static represetation.
Tversky, B., Morrison, J. B., & Betrancourt, M. (2002). Animation: can it facilitate? Int. J. Human-Computer Studies Schnotz & Kulhavy, 57, 247–262. http://doi.org/10.1006/ijhc.1017

**Client-side interactions**
Client-side interactions produce an html rendering of the notebook that can be shared with those that do not have R or the underlying data. The main drawback with client-side interactions is that they support only simple annotation, zooming, and linking and not any data filtering or transformation.

Useful client-side interaction packages include:
* DataTable and other HTML widgets (See Chapter 10 for more details)
* ggiraph provides simple annotation and linking
* ggplotly provides zooming based on brushed data, as well as more flexible annotation and linking


**Server-side interactions**
Server-sider interactions provide great flexibility in linking graphs and selecting, filtering and transforming data based on user input.

Useful approaches for server-side interactions
* Shiny notebooks provides simple way of adding data selection and linking
* Shiny applications provides may layout options that notebooks don't support


```{r, message=FALSE, warning=FALSE}
library(DT)
library(data.table)
library(sparkline)
library(ggiraph)
#devtools::install_github("thomasp85/patchwork")
library(patchwork) # For combining multiple plots with ggiraph
library(cowplot)

library(plotly)
library(sf) # simple features for maps
# devtools::install_github("hrbrmstr/albersusa")

library(albersusa)
library(gapminder)
library(tidyverse)

#library(crosstalk)

```


##  DT: Interactive tables
DT is a very powerful package for creating interactive tables. By default it supports sorting sorting and searching. It can also be configured so that each column can be used to filter the data.  Extensions even make it possible to fill the cells with small graphical summaries--sparklines. 
Kowarik, A., Meindl, B., & Templ, M. (2014). sparkTable: Generating graphical tables for websites and documents with R. R J., 7(1), 24-37.

```{r}

mtcars %>% select(mpg, cyl, wt, hp, qsec) %>% 
  datatable(filter = 'top')

```

##  DT: Interactive tables with sparklines
```{r, echo=FALSE, warning=FALSE}
## https://leonawicz.github.io/HtmlWidgetExamples/ex_dt_sparkline.html

l.iris= iris %>% gather(key = Var, Value, -Species)
r = range(l.iris$Value) # To place boxplots on an absolute range

# Set parameters of sparklines --------------------------------------------
bar_string <- "type: 'bar', barColor: 'orange', negBarColor: 'purple', highlightColor: 'black'"
cb_bar = JS(paste0("function (oSettings, json) { $('.spark:not(:has(canvas))').sparkline('html', { ", bar_string, " }); }"), collapse="")

line_string <- "type: 'line', lineColor: 'black', width: '200', fillColor: '#ccc', highlightLineColor: 'orange', highlightSpotColor: 'orange'"

line_string2 = "type: 'line', lineColor: 'black', width: '200', fillColor: '#ccc', highlightLineColor: 'orange', highlightSpotColor: 'orange', chartRangeMin: ', r[1],' ,chartRangeMax: ', r[2]'"
cb_line = JS(paste0("function (oSettings, json) { $('.spark:not(:has(canvas))').sparkline('html', { ", line_string, ", chartRangeMin: ", r[1], ", chartRangeMax: ", r[2], " }); }"), collapse="")

box_string <- "type: 'box', lineColor: 'black',  width: '300', whiskerColor: 'black', outlierFillColor: 'black', outlierLineColor: 'black', medianColor: 'black', boxFillColor: 'orange', boxLineColor: 'black'"
cb_box1 = JS(paste0("function (oSettings, json) { $('.spark:not(:has(canvas))').sparkline('html', { ", box_string," }); }"), collapse="")

cb_box2 = JS(paste0("function (oSettings, json) { $('.spark:not(:has(canvas))').sparkline('html', { ", box_string,", chartRangeMin: ", r[1], ", chartRangeMax: ", r[2], " }); }"), collapse="")


s.iris <- l.iris %>% 
  group_by(Species, Var) %>% 
  summarise(Mean = round(mean(Value), 1), SD = round(sd(Value),2), Min = min(Value), Max = max(Value), 
    Samples = paste(Value, collapse = ","), Series = paste(Value, collapse = ",")) %>% # Specifies spark data based on individual values
  arrange(Var) %>% 
  data.table

# Set column and callback definitions  --------------------------------------------
# "targets"" specify column for sparklines
cd <- list(list(targets = 6, render = JS("function(data, type, full){ return '<span class=sparkSamples>' + data + '</span>' }")), 
           list(targets = 7, render = JS("function(data, type, full){ return '<span class=sparkSeries>' + data + '</span>' }")))

cb = JS(paste0("function (oSettings, json) {\n  $('.sparkSeries:not(:has(canvas))').sparkline('html', { ", 
               line_string,"});\n  $('.sparkSamples:not(:has(canvas))').sparkline('html', { ", 
               box_string,", chartRangeMin: ", r[1], ", chartRangeMax: ", r[2], " });\n}"), collapse = "")
      # r defined by range of the data 

# Render data table  --------------------------------------------
iris.dt <- datatable(data.table(s.iris), rownames = FALSE, 
                      options = list(pageLength = 12, lengthMenu = c(6, 12, 24),
                      columnDefs = cd, fnDrawCallback = cb, autoWidth = TRUE, 
                      columnDefs = list(list(width = '100px', targets = c(6, 7)))
                ))

iris.dt$dependencies <- append(iris.dt$dependencies, htmlwidgets:::getDependency("sparkline"))
iris.dt

```



##  ggiraph: Simple annotation
ggirpah makes it very easy to create interacive graphic.  This package supports three interactive elements
tooltip: mouse-over tooltips display information when mouse is over elements.
data_id: highlights elements that share the same variable value.
onclick: executes java script when elements are clicked, such as opening a webpage associated with the point.



Annotate details on hover
```{r, warning=FALSE, message=FALSE}

mtcars.df = mtcars
mtcars.df$name = rownames(mtcars.df)

mpg.plot =  ggplot(mtcars.df, aes( x = disp, y = mpg))+  
  geom_point_interactive(aes(tooltip = name), size = 2) +  	
  labs(title = "Hover to reveal name")
	
ggiraph(code = 	print(mpg.plot))

```

## ggiraph: Highlighting and multi-plot linking

```{r, linked-plot, warning=FALSE, message=FALSE}

hp.plot = ggplot(mtcars.df) +
  geom_point_interactive(aes(x = hp, y = mpg,
    tooltip = name, data_id = name), size = 2) +
  labs(title = "Hover to locate vehicle in both plots")

wt.plot <- ggplot(mtcars.df) +
  geom_point_interactive(aes(x = wt, y = mpg,
    tooltip = name, data_id = name), size = 2)

girafe(code = print(hp.plot + wt.plot))



gapminder.df = gapminder %>% group_by(continent, year) %>% 
  summarise(m.gdpPercap = mean(gdpPercap), m.lifeExp = mean(lifeExp))

gap_scatter.plot = 
ggplot(gapminder.df) + 
  geom_point_interactive(aes(x=m.gdpPercap, y=m.lifeExp, colour = continent,
    tool_tip = continent, data_id = continent), alpha = .8) +
  labs(title = "Hover to highlight data for a continent") +
  theme(legend.position = "bottom")

 
gap_timeline.plot = 
ggplot(gapminder.df) + 
  geom_line_interactive(aes(x=year, y=m.lifeExp, group = continent, colour= continent,
      tool_tip = continent, data_id = continent),alpha = .8) +
  theme(legend.position = "none")
 
girafe(code = print(gap_scatter.plot + gap_timeline.plot)) %>% 
   girafe_options(opts_hover(css = "stroke-width:3px;"))
 
```



## plotly: Map annotation with flexible tooltip annotation
Plotly supports filtering, highlighting, and linking views
Legend selection of data
Brush zooming
Slider filter and animation

ggplotly provides more flexibiltiy with tooltip annotations compared to ggiraph

```{r, interactive_map, warning=FALSE}
# 
# us_laea <- usa_sf("laea")
# us_laea = us_laea %>% mutate(density_2014 = pop_2014/census_area)
# 
# map.plot =  
# ggplot(us_laea) + 
#   geom_sf(aes(fill = log(density_2014), 
#               text = paste(name, " density: ", round(density_2014, 0),
#                            "people per square mile")))+
#   scale_fill_viridis_c(option="magma") +
#   labs(title = "Hover show details") +
#   theme_void()

# ggplotly(map.plot, tooltip = "text") %>%
#   style(hoverlabel = list(bgcolor = "grey85"), hoveron = "fill")

```

## plotly: Detailed annotation, legend-based selection, and brush to zoom



```{r, warning=FALSE, message=FALSE}

gap.plot = 
  ggplot(gapminder.df, 
         aes(x = m.gdpPercap/1000, y = m.lifeExp, colour = continent, text = year)) +
  geom_point() +
  labs(title = "Brush points to zoom in on a subset of data")

ggplotly(gap.plot, tooltip = c("text", "x", "y", "colour")) 

```




## plotly: Linked plots
Unlike the linked plot with ggiraph, there is no easy way to avoid showing a duplicate legend.

```{r, linked_plot, message=FALSE, warning=FALSE}

gapminder.sd <- highlight_key(gapminder.df, ~continent)

p1 = 
ggplot(gapminder.sd, aes(m.gdpPercap/1000, m.lifeExp, colour = continent)) +
  geom_point() +
  labs(title = "Hover show details, click to highlight group")+
  theme(legend.position = "none")

p2 = 
ggplot(gapminder.sd, aes(year, m.lifeExp, colour = continent)) +
  geom_point()+
  theme(legend.position = "bottom")

pl1 = ggplotly(p1, tooltip = c("continent", "m.gdpPercap"))
pl2 = ggplotly(p2, tooltip = c("continent", "year"))

subplot(pl1, pl2, nrows = 2)

```



## plotly: Highlight and annotate data and fit
Adding year as grouping factor makes it possible to include in tooltip annotation
Highlight commands makes all but selected items semi-transparent
Highlight from legend or points
```{r, warning = FALSE, message=FALSE}

gap.sd = highlight_key(gapminder, ~new) # Annotates points
gap.sd = highlight_key(gapminder, ~continent) # Annotates fits

gap.plot =
ggplot(gap.sd, aes(gdpPercap/1000, lifeExp, colour = continent)) +
  geom_point(aes(text = country, group = year), size =.75) +
  geom_smooth(se = FALSE, method = "lm") +
  scale_x_log10() +
  labs(title = "Hover to highlight and show details")
  
ggplotly(gap.plot,  tooltip = c("text", "y", "year"))%>% 
  highlight("plotly_hover") 

```


## plotly: Time series scatterplot animation

ids and frame specify the data elements for the animation and slider
```{r, scatterplot_time, warning=FALSE}

time.plot = 
ggplot(gapminder, aes(gdpPercap/1000, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10() +
  labs(title = "Slide to change year or Play to animate")

ggplotly(time.plot)

```
  
  